---
title: "Radiomics for anal cancer prognosis"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, results = 'hide', warning = FALSE, message = FALSE}
library(tidyverse)
library(survival)
library(survcomp)
library(broom)
library(caret)
library(mRMRe)

Sys.setenv(KMP_DUPLICATE_LIB_OK=TRUE) # when installing the new mRMRe from GitHub, libomp is loaded twice,
                                      # crashing R, probably related to my installation of clang-omp.

clinical <- read_csv("../data/clinical_clean.csv")
radiomics <- read_csv("../data/features_all.csv") %>% 
  select(-contains("diagnostics"), -contains("Image"), -contains("Mask"), -contains("lbp")) %>%
  rename_at(vars(-patient_id), .funs = ~paste("radiomics_", ., sep = ""))
inception <- read_csv("../data/inception_features.csv") %>% 
  select(-contains("Image"), -contains("Mask"))

data <- inner_join(clinical, radiomics) %>% inner_join(inception)
clinical_cols <- 
  list(
    clinical_basic_sex_f  = sym("SEX_F"),
    clinical_basic_age    = sym("AGE"),
    clinical_basic_tstage = sym("cT"),
    clinical_basic_nstage = sym("cN"),
    clinical_basic_chemo  = sym("Concurrent chemo (1=yes, 0=no)"),
    clinical_basic_volume = sym("GTV Size (cc)"),
    clinical_basic_hiv    = sym("HIV (0=no, 1=yes)"),
    clinical_grade        = sym("Hist Grade"),
    clinical_size         = sym("Size"),
    # clinical_dose         = sym("Primary Dose (cGy)"),
    clinical_hb           = sym("Hb"),
    clinical_hb_100       = sym("Hb >100 (1=yes, 0=no)"),
    clinical_wbc          = sym("WBC"),
    clinical_plt          = sym("Plt"),
    clinical_neut         = sym("Neut"),
    clinical_lymph        = sym("Lymph"),
    clinical_plr          = sym("Platelets:Lymphocyte ratio"),
    clinical_nlr          = sym("Neutrophil:Lymphocyte ratio"),
    clinical_nlr2         = sym("NLR>2 (1= >2, 0=<2)"))

data <- data %>%
  rename(!!!clinical_cols) %>%
  filter(patient_id != 994) %>%
  drop_na(contains("clinical"))
```


## Overview
- `r nrow(data)` anal cancer cases
- `r data %>% select(contains("clinical")) %>% ncol` clinical variables
- `r data %>% select(contains("radiomics")) %>% ncol` radiomics features extracted:
  - original image + filters available in Pyradiomics: wavelet, square, square root, exponential, gradient
  - 2 scales (spacing = .81 mm and 1.62 mm)
- image preprocessing:
  - resampling to uniform spacing (either .81 or 1.62 in x and y, 3.75 mm in z)
  - quantization to fixed bin count = 64 (as recommended by IBSI)
- 1000 80%/20% train-test splits
- for each split:
  - train a "basic" clinical model with 7 variables:
    - age
    - sex
    - T stage
    - N stage
    - concurrent chemotherapy
    - primary volume
    - HIV
  - select 9 clinical and 9 radiomics variables using mRMRe (minimum redundancy-maximum relevance)
  - select 9 best variables from clinical and radiomics
  - fit Cox proportional hazards model for OS, DFS, DM
  - fit Fine-Gray model for LF and LRF with death as competing risk
  - compute concordance index on the held-out set

## Count missing times/events
```{r}
clinical %>% 
  select(contains("time")) %>%
  summarise_all(funs(sum(is.na(.))))
```

```{r}
clinical %>% 
  select(contains("event")) %>%
  summarise_all(funs(sum(is.na(.))))
```


## Calculate event rates for all outcomes of interest
```{r}
clinical %>% 
  select(contains("event")) %>%
  summarise_all(mean, na.rm=T) %>%
  gather(event, value) %>%
  ggplot(aes(x = event, y = value)) +
  geom_col() 
```

```{r}
mrmr_select <- function(data, feature_count = 10, fixed_feature_count = 0) {
  data_df <- as.data.frame(data)
  data_df$surv <- Surv(data_df$time_col, data_df$event_col)
  data_df$time_col <- NULL
  data_df$event_col <- NULL
  mrmr_data <- mRMR.data(data_df)
  mrmr_obj <- mRMR.ensemble(data = mrmr_data,
                            target_indices = c(length(data_df)),
                            solution_count = 1,
                            feature_count = feature_count)
  return(featureNames(mrmr_obj)[solutions(mrmr_obj)[[1]]])
}

run_cox_model <- function(train, test, feature_count = 10, fixed_feature_count = 0) {

    selected_features <- mrmr_select(train, feature_count, fixed_feature_count)

    model <- train %>%
      select(selected_features, time_col, event_col) %>%
      coxph(Surv(time_col, event_col) ~ ., data=.)
    
    predicted <- predict(model,
                         test %>% select(selected_features),
                         type = "risk")

    # print("predicted:")
    # print(predicted)
    # print("time:")
    # print(test$time_col)
    # print("event:")
    # print(test$event_col)
    # print(paste(rep("=", 20)))
    return(tidy(model, exponentiate = TRUE) %>%
             mutate(ci=concordance.index(predicted,
                                         test$time_col,
                                         test$event_col,
                                         outx = FALSE)$c.index))
}

run_finegray_model <- function(train, test, feature_count = 10, fixed_feature_count = 0) {

    selected_features <- mrmr_select(train, feature_count, fixed_feature_count)
    
    pdata <- train %>% 
      rowwise %>%
      mutate(ftime = min(time_col, competing_time_col),
             fstatus = ifelse(event_col == 1, 1, competing_event_col * 2)) %>%
      ungroup %>%
      select(selected_features, ftime, fstatus, -time_col, -competing_time_col, -event_col, -competing_event_col) %>%
      finegray(Surv(ftime, factor(fstatus, levels=c(0, 1, 2))) ~ ., data=.)

    
    model <- coxph(Surv(fgstart, fgstop, fgstatus) ~ . - fgwt, weights = fgwt, data=pdata)
    
    predicted <- predict(model,
                         test %>% select(selected_features) %>% mutate(fgwt = rep(1, nrow(test))),
                         type = "risk")

    return(tidy(model, exponentiate = TRUE) %>%
             mutate(ci=concordance.index(predicted,
                                         test$time_col,
                                         test$event_col,
                                         outx = FALSE)$c.index))
}

find_optimal_feature_count <- function(data, lower = 5, upper = 20, fixed_feature_count = 0) {
  folds <- createFolds(as.factor(data$event_col), k = 3)
  results <- tibble(n_feats = numeric(), fold = numeric(), ci = numeric())
  for(n_feats in lower:upper) {
    for(i in 1:length(folds))  {
      train_fold <- data[-folds[[i]],]
      test_fold <- data[folds[[i]],]
      ci <- run_cox_model(train_fold, test_fold, n_feats, fixed_feature_count = min(n_feats - 1, fixed_feature_count))$ci[[1]]
      results <- results %>% add_row(n_feats = n_feats,
                                     fold = i,
                                     ci = ci)
    }
  }
  top_n_feats <- results %>% group_by(n_feats) %>% summarise(ci=mean(ci)) %>% top_n(1) %>% select(n_feats)
  return(top_n_feats[[1]])
}

run_analysis <- function(data, feature_count = 10 , n_iter = 100, fixed_feature_count = 0, train_radiomics_only = TRUE) {
  results <- tibble(
    ci = numeric(),
    variable = character(),
    iteration = numeric(),
    model = character())
  
  if("competing_event_col" %in% colnames(data)) {
    model_fn <- run_finegray_model
    target_cols <- c("time_col", "competing_time_col", "event_col", "competing_event_col")
  } else {
    model_fn <- run_cox_model
    target_cols <- c("time_col", "event_col")
  }

  for(i in 1:n_iter) {
    train_indices <- createDataPartition(as.factor(data$event_col), p = .8, list = FALSE)
    train <- data[train_indices,]
    test <- data[-train_indices,]
    
    if(feature_count == "optim") {
      feature_count_clin <- train %>%
        select(contains("clinical_basic"), time_col, event_col) %>%
        find_optimal_feature_count(upper = ncol(.) - 2)
    } else {
      feature_count_clin <- min(ncol(data %>%
                                       select(contains("clinical_basic"))), 
                                feature_count)
    }
    
    clinical_basic_model <- model_fn(train %>% select(contains("clinical_basic"), target_cols),
                                     test,
                                     feature_count_clin,
                                     fixed_feature_count = 0)
    selected_clinical_basic <- gsub("`", "", clinical_basic_model$term)
    
    if(feature_count == "optim") {
      feature_count_rad <- train %>%
        select(contains("radiomics"), time_col, event_col) %>%
        find_optimal_feature_count(upper = ncol(train %>% select(contains("clinical"))))
    } else {
      feature_count_rad <- feature_count
    }
    
    if(train_radiomics_only) {
      radiomics_model <- model_fn(train %>% select(contains("radiomics"), target_cols),
                                  test,
                                  feature_count_rad)
      selected_radiomics <- gsub("`", "", radiomics_model$term)
    }

    if(feature_count == "optim") {
      feature_count_comb <- train %>%
        select(selected_clinical_basic, selected_radiomics, time_col, event_col) %>%
        find_optimal_feature_count(upper = ncol(.) - 2)
    } else {
      feature_count_comb <- feature_count
    }
    
    combined_model <- model_fn(train %>% select(selected_clinical_basic, contains("radiomics"), target_cols),
                               test,
                               feature_count = feature_count_comb,
                               fixed_feature_count = min(fixed_feature_count, feature_count_comb))
    selected_combined <- gsub("`", "", combined_model$term)
    
    if(feature_count == "optim") {
      feature_count_clin_adv <- train %>%
        select(contains("clinical"), time_col, event_col) %>%
        find_optimal_feature_count(upper = ncol(.) - 2)
    } else {
      feature_count_clin_adv <- feature_count
    }
    
    clinical_adv_model <- model_fn(train %>% select(contains("clinical"), target_cols),
                                   test,
                                   feature_count_clin_adv,
                                   fixed_feature_count = min(fixed_feature_count, feature_count_clin_adv))
    selected_clinical_adv <- gsub("`", "", clinical_adv_model$term)
    
    # inception_model  <- run_cox_model(train %>% select(contains("inception"),
    #                                                    time_col,
    #                                                    event_col),
    #                                   test,
    #                                   feature_count_rad)
    # selected_inception <- gsub("`", "", inception_model$term)
    
    for(v in selected_clinical_basic) {
      results <- results %>% add_row(
        ci = clinical_basic_model$ci[[1]],
        variable = v,
        iteration = i,
        model = "clinical_basic")
    }
    
    if(train_radiomics_only) {
      for(v in selected_radiomics) {
        results <- results %>% add_row(
          ci = radiomics_model$ci[[1]],
          variable = v,
          iteration = i,
          model = "radiomics")
      }
    }
    
    for(v in selected_combined) {
      results <- results %>% add_row(
        ci = combined_model$ci[[1]],
        variable = v,
        iteration = i,
        model = "combined")
    }
    
    for(v in selected_clinical_adv) {
      results <- results %>% add_row(
        ci = clinical_adv_model$ci[[1]],
        variable = v,
        iteration = i,
        model = "clinical_adv")
    }
    
    # for(v in selected_inception) {
    #   results <- results %>% add_row(
    #     ci = inception_model$ci[[1]],
    #     variable = v,
    #     iteration = i,
    #     model = "inception")
    # }
  }
  return(results)
}

plot_variable_frequencies <- function(data) {
  plt <- data %>% 
    group_by(model, variable) %>% 
    summarise(percent_selected = n_distinct(iteration) / 1000) %>%
    top_n(10, wt = percent_selected) %>%
    ungroup() %>% 
    arrange(model, percent_selected) %>%
    mutate(order = row_number())

  p <- plt %>%
    ggplot(aes(x = order, y = percent_selected)) +
    geom_col() +
    facet_grid(rows = vars(model), scales = "free_y") + 
    scale_x_continuous(
      breaks = plt$order,
      labels = plt$variable,
      expand = c(0,0)) + 
    coord_flip()
  return(p)
}
```

## Overall survival

```{r}
fixed_feature_count <- data %>% select(contains("clinical_basic")) %>% ncol
```

```{r, results = 'hide', warning = FALSE, message = FALSE}
results_os <- run_analysis(data %>% 
                             select(contains("clinical"),
                                    contains("radiomics"),
                                    time_col = death_time, 
                                    event_col = event_os),
                           n_iter = 1000, 
                           feature_count = 9,
                           fixed_feature_count = fixed_feature_count)
```

### CI distributions across random splits
```{r}
results_os %>%
  ggplot(aes(x = model, y = ci)) +
  geom_boxplot()
```

### Variable selection frequencies across random splits for each model
```{r}
results_os %>% plot_variable_frequencies()
```

### Results summary for overall survival
```{r}
results_os %>%
  group_by(model) %>%
  summarise_at(vars(ci), list(mean_ci=mean, median_ci=median, sd_ci=sd))
```

## Disease-free survival

```{r, results = 'hide', warning = FALSE, message = FALSE}
results_dfs <- run_analysis(data %>%
                              select(contains("clinical"), 
                                     contains("radiomics"), 
                                     time_col = dfs_time, 
                                     event_col = event_dfs),
                            n_iter = 1000, 
                            feature_count = 9, 
                            fixed_feature_count = fixed_feature_count)
```

### CI distributions across random splits
```{r}
results_dfs %>%
  ggplot(aes(x = model, y = ci)) +
  geom_boxplot()
```

### Variable selection frequencies across random splits for each model
```{r}
results_dfs %>% plot_variable_frequencies()
```

### Results summary for disease-free survival
```{r}
results_dfs %>%
  group_by(model) %>%
  summarise_at(vars(ci), list(mean_ci=mean, median_ci=median, sd_ci=sd))
```

## Local failure

```{r, results = 'hide', warning = FALSE, message = FALSE}
results_lf <- run_analysis(data %>%
                             select(contains("clinical"),
                                    contains("radiomics"),
                                    time_col = lf_time, 
                                    event_col = event_lf,
                                    competing_time_col = death_time,
                                    competing_event_col = event_os),
                           n_iter = 1000,
                           feature_count = 9,
                           fixed_feature_count = fixed_feature_count,
                           train_radiomics_only = FALSE) # radiomics-only model causes numerical error, need to investigate further
```

### CI distributions across random splits
```{r}
results_lf %>%
  ggplot(aes(x = model, y = ci)) +
  geom_boxplot()
```

### Variable selection frequencies across random splits for each model
```{r}
results_lf %>% plot_variable_frequencies()
```

### Results summary for local failure
```{r}
results_lf %>%
  group_by(model) %>%
  summarise_at(vars(ci), list(mean_ci=mean, median_ci=median, sd_ci=sd))
```

## Local/regional failure

```{r, results = 'hide', warning = FALSE, message = FALSE}
results_lrf <- run_analysis(data %>%
                              select(contains("clinical"),
                                     contains("radiomics"), 
                                     time_col = lrf_time, 
                                     event_col = event_lrf),
                            n_iter = 1000, 
                            feature_count = 9,
                            fixed_feature_count = fixed_feature_count)
```

### CI distributions across random splits
```{r}
results_lrf %>%
  ggplot(aes(x = model, y = ci)) +
  geom_boxplot()
```

### Variable selection frequencies across random splits for each model
```{r}
results_lrf %>% plot_variable_frequencies()
```

### Results summary for local/regional failure
```{r}
results_lrf %>%
  group_by(model) %>%
  summarise_at(vars(ci), list(mean_ci=mean, median_ci=median, sd_ci=sd))
```

## Distant metastasis

```{r, results = 'hide', warning = FALSE, message = FALSE}
results_dm <- run_analysis(data %>%
                             select(contains("clinical"),
                                    contains("radiomics"),
                                    time_col = dm_time, 
                                    event_col = event_dm),
                           n_iter = 1000,
                           feature_count = 9,
                           fixed_feature_count = fixed_feature_count)
```

### CI distributions across random splits
```{r}
results_dm %>%
  ggplot(aes(x = model, y = ci)) +
  geom_boxplot()
```
### Variable selection frequencies across random splits for each model
```{r}
results_dm %>% plot_variable_frequencies()
```

### Results summary for distant metastasis
```{r}
results_dm %>%
  group_by(model) %>%
  summarise_at(vars(ci), list(mean_ci=mean, median_ci=median, sd_ci=sd))
```